---
title: "Who checks the museum names?"
subtitle: "America's most duplicated museum name, and five collisions that looked too good to check."
author: "J.P. Branson"
date: 2026-10-05
slug: duplicate-museum-names
categories: [data]
tags: [names, museums]
draft: true
output:
  blogdown::html_page:
    toc: false
    fig_caption: true
---

```{r setup, include=FALSE}
dn_root <- "."
source("_setup.R", local = TRUE)
knitr::opts_chunk$set(fig.path = "figures/")
num <- function(x) format(x, big.mark = ",")
word <- function(n) c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten", "eleven", "twelve")[n]
Word <- function(n) tools::toTitleCase(word(n))
link <- function(text, url) sprintf("[%s](%s)", text, url)
cp <- dn_payload("checkpoint.csv")
top <- dn_payload("headline.csv")
jail <- dn_payload("headline_members.csv")
groups <- dn_payload("collision_groups.csv")
records <- dn_payload("collision_members.csv")
chains <- dn_payload("chains.csv")
runner_up <- dn_payload("runner_up_names.csv")
seed <- dn_payload("seed_records.csv")
bangor <- seed[tolower(seed$name_raw) == "international cryptozoology museum", ]
chain_n <- function(id) chains$n_locations[chains$chain_id == id]
# The prose below names these results; stop the knit if a re-export changes them.
stopifnot(
  top$display_name == "Old Jail Museum", top$museums == nrow(jail), top$museums > top$runner_up_museums,
  sum(grepl("Historical", jail$operator)) == top$museums - 1L, top$chain_locations == 3L,
  identical(groups$status[match(c("100th meridian museum", "billy the kid museum"), groups$group)],
            c("confirmed", "unresolved_record")),
  sum(groups$status == "not_a_collision") == 3L, nrow(records) == 12L, sum(groups$counting) == 7L,
  nrow(bangor) == 4L, sum(bangor$counted) == 1L,
  abs(mean(grepl("historical society", runner_up$name_expanded)) - 0.5) < 0.1
)
```

The International Cryptozoology Museum in Bangor, Maine, bills itself as
["The World's Only Cryptozoology Museum"](https://cryptozoologymuseum.com/).
Standing in front of it, the question that nags is not whether Bigfoot is real.
It is: who checks?

Nobody does. There is no registrar of museum names. Anyone can open The Museum of
Anything, call it National or International, and no office writes back to say the
name is taken. So I checked. Not the world, just the United States, using two public
lists of museums and a fair amount of tidying.

## The most duplicated museum name in America

Start with the easy version of the question: which name do the most museums share?

After merging records that describe the same place and setting chains aside, the
lists give me `r num(cp$non_chain_institutions)` museums, or at least
`r num(cp$non_chain_institutions)` things the lists call museums, under
`r num(cp$distinct_names)` different names. Most are safe:
`r num(cp$names_used_once)` of those names belong to exactly one museum. The rest
get crowded quickly.

```{r name-ladder}
ladder <- dn_payload("name_ladder.csv")
ladder <- merge(data.frame(museums_sharing = seq_len(max(ladder$museums_sharing))), ladder, all.x = TRUE)
ladder$names[is.na(ladder$names)] <- 0
shared_by <- paste(ladder$museums_sharing, ifelse(ladder$museums_sharing == 1, "museum", "museums"))
knitr::kable(data.frame(shared_by, num(ladder$names)), row.names = FALSE, align = "lr",
             col.names = c("A name shared by", "How many names"))
```

The one name at the top is **`r top$display_name`**.

`r Word(top$museums)` unrelated museums, each in a former jail, each run by a
different local group: `r word(top$museums - 1)` by a historical society or
council, one by a family.

```{r old-jail}
jail$operator <- sub(" \\(.*\\)$", "", jail$operator)
jail$operator[grepl("^Family-owned", jail$operator)] <- "The McBride family"
knitr::kable(data.frame(paste0(jail$place, ", ", jail$state), link(jail$operator, jail$evidence_url)),
             row.names = FALSE, col.names = c("Old Jail Museum in", "Run by"))
```

It is a very reasonable name. The county builds a new jail. The old one is too
sturdy to knock down and too grim for offices, so the historical society moves in
and the sign more or less writes itself. Nobody in Thompson Falls, Montana, had to
ask Hayesville, North Carolina, for permission.

Second place is a `r top$runner_up_names`-way tie at `r word(top$runner_up_museums)`.
About half are historical societies named for a popular county or town (Franklin
County, Greene County, Jackson County); most of the rest are plain descriptions
such as Pioneer Museum, Telephone Museum and Dinosaur Museum. I did not vet all
`r top$runner_up_names`. I did check that none of them can catch up through spelling
variants, and a few are probably one organization with several branches, which would
only push them down.

The lead grows if you loosen the rule. I counted exact names, so the Monticello Old
Jail Museum and the Old Jail Museum Complex are different names. Count anything with
"old jail" in it and the lists have `r top$related_museums`.

`r Word(top$chain_locations)` more Old Jail Museums are missing from the
`r word(top$museums)` on purpose. Each belongs to an operator that runs other
museums too (in St. Augustine, Florida; Allegan, Michigan; and Sandersville,
Georgia), and I count shared operators and chains separately. That rule matters more
elsewhere: the names that repeat most in the raw lists are brands, led by Play
Street Museum (`r chain_n("play_street")` locations), Museum of Illusions
(`r chain_n("museum_of_illusions_global")`) and Ripley's (`r chain_n("ripleys")`).
A brand repeating is a business plan, not a coincidence.

## Five collisions that looked too good to check

Old jails are a category. The better collisions are names specific enough that you
would bet there is only one. I went looking for those: names with no town, county or
generic word in them, shared by museums a long way apart. I picked
`r word(nrow(groups))`, then checked every record behind them against the best
public source I could find, usually the museum's own website or an official tourism
listing. Is it a real museum the public can visit? Does it appear to be operating?
Is it unrelated to its namesake?

```{r collisions, fig.asp=0.62, fig.alt="Grouped bar chart of five museum names, comparing records in the data with museums that passed the check. Billy the Kid Museum: 3 records, 2 museums. 100th Meridian Museum: 2 and 2. Santa Claus Museum: 2 and 1. The Mermaid Museum: 2 and 1. Salt and Pepper Shaker Museum: 3 and 1.", fig.cap="Twelve records, seven museums, two real collisions. Checked October 5, 2026."}
dat <- groups[order(groups$counting, groups$records), ]
dat$display_name <- sub("^Mermaid", "The Mermaid", dat$display_name)
dat$display_name <- factor(dat$display_name, levels = dat$display_name)
series <- c("Records in the data", "Museums that passed the check")
long <- rbind(data.frame(name = dat$display_name, series = series[1], n = dat$records),
              data.frame(name = dat$display_name, series = series[2], n = dat$counting))
long$series <- factor(long$series, levels = series)
dodge <- position_dodge2(width = 0.7, reverse = TRUE, padding = 0.12)
ggplot(long, aes(n, name, fill = series)) +
  geom_col(position = dodge, width = 0.7) +
  geom_text(aes(label = n), position = position_dodge2(width = 0.7, reverse = TRUE),
            hjust = -0.7, size = 4.1, colour = dn_ink$primary) +
  scale_fill_dn() +
  scale_x_continuous(limits = c(0, max(long$n) + 0.4), breaks = 0:max(long$n), expand = c(0, 0)) +
  theme_dupnames(base_size = 14) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(colour = dn_ink$grid, linewidth = 0.3),
        axis.line.x = element_blank(), axis.text.y = element_text(colour = dn_ink$primary)) +
  labs(x = NULL, y = NULL, title = "What the lists said, and what survived",
       subtitle = "Two museums left on a bar means the collision is real.",
       caption = dn_caption(c("overture", "imls"), "Checks by the author against public sources."))
```

**100th Meridian Museum: real.** One is in
[Cozad, Nebraska](https://visitnebraska.com/cozad/100th-meridian-museum), run by the
Cozad Historical Society. The other is in
[Erick, Oklahoma](https://www.travelok.com/listings/view.profile/id.37151), about 390
miles south, run by its own nonprofit. Cozad has the older claim to the line.
[Union Pacific](https://www.up.com/aboutup/train_town/cozad_ne/index.htm) says its
railroad reached the 100th meridian there in 1866, and that the town's namesake,
John J. Cozad, spotted the 100th Meridian sign from a train in 1872. Erick is, in
the
[Oklahoma Historical Society](https://www.okhistory.org/publications/enc/entry.php?entry=ER001)'s
words, "six miles from the Oklahoma-Texas border", and that stretch of border is
drawn along the meridian.

**Billy the Kid Museum: real, with a ghost.** The
[Fort Sumner, New Mexico](https://www.billythekidmuseumfortsumner.com/), museum has
been run by the Sweet family since 1953. The
[Hico, Texas](https://www.billythekidmuseumhico.com/), museum is a separate
nonprofit. They share a name and disagree about the ending. Fort Sumner's
[state tourism listing](https://www.newmexico.org/places-to-visit/regions/southeast/billy-the-kid-museum-in-fort-sumner/)
says "we firmly believe that he was in fact shot and Killed by Pat Garrett on July 14
1881". Hico's museum says the history books got it wrong and the Kid "actually
survived his famous gun fight and hid in Hico, Texas the last seventeen years of his
life". The lists also had a third Billy the Kid Museum, in Clovis, New Mexico. Its
coordinates match an address that belongs to a
[tire shop](https://local.yahoo.com/tires/nm/clovis/). The same phantom sits on a
[road-trip site](https://maps.roadtrippers.com/us/clovis-nm/attractions/billy-the-kid-museum-clovis-nm)
with Hico's web address and a one-star review: "The museum is awesome but is NOT in
Clovis."

**Santa Claus Museum: a near miss.** The one in
[Columbus, Texas](https://santamuseum.org/about/), holds
[more than 5,000 Santas](https://www.columbushpt.com/index.php/historical-properties/santa-claus-museum/)
and says there are "only two museums dedicated to Santa Claus in the entire United
States". The other is in Santa Claus, Indiana, and it calls itself the
[Santa Claus Museum & Village](https://santaclausmuseum.org/). A 2018 federal list
has it as plain Santa Claus Museum and a newer map has the longer name, so my data
counted one museum twice and then matched its old name to Texas. Two museums, two
names, three records.

**The Mermaid Museum: one museum and one marketing campaign.**
[Berlin, Maryland](https://www.berlinmermaidmuseum.com/), has a real one. The record
in Hollywood sits about a block from a studio that hosted the
[POPSUGAR x Freeform Mermaid Museum](https://www.timeout.com/los-angeles/things-to-do/popsugar-x-freeform-mermaid-museum),
a promotional pop-up that ran for four days in March 2018. I found no mermaid museum
there now. My best guess is that the pop-up never left the map. (Washington state
has an International Mermaid Museum.
[It opened two days after Berlin's](https://hyperallergic.com/the-curious-emergence-of-two-mermaid-museums-in-the-us),
and both founders were surprised.)

**Salt and Pepper Shaker Museum: one museum, twice, and a gallery.** The
[Gatlinburg, Tennessee](https://thesaltandpeppershakermuseum.com/), museum calls
itself "The World's only Salt and Pepper Shaker Museum", with a footnote: "Well, we
do have a Sister Museum in Spain". The second record is the same museum from the
2018 list, with coordinates in San Francisco. The third is in
[Traer, Iowa](https://www.traer.com/museums). It is real, the city owns it, and it
has "more than 15,000 sets of shakers". It is also called the Salt & Pepper Shaker
Gallery. Gatlinburg's claim survives by one word.

So: `r word(nrow(records))` records, `r word(sum(groups$counting))` museums, and
`r word(sum(groups$counting >= 2))` real collisions out of `r word(nrow(groups))`.
Names nobody coordinates collide by accident. Lists nobody curates invent collisions
that are not there. In this batch, the second kind was more common.

## Back to Bangor

The International Cryptozoology Museum appears `r word(nrow(bangor))` times in my
data: once at its current address and `r word(sum(!bangor$counted))` times in
Portland, Maine, which it has left. One museum, and no other museum in these lists
shares its name.

Whether it is the world's only is a different claim, and nobody checks that one
either. The same lists have a Cryptozoology & Paranormal Museum in Littleton, North
Carolina, which WALTER magazine
[profiled in 2023](https://waltermagazine.com/explore/its-cryptic-the-bigfoot-museum-in-littleton/)
as the Bigfoot museum. I am not refereeing.

Next: churches, which noticed this problem a few centuries ago and solved it by
counting. First Baptist, Second Baptist, Third.

## How this was counted

- **Sources.** [Overture Maps](https://overturemaps.org/) places, release 2026-08-19,
  and the Institute of Museum and Library Services'
  [2018 Museum Universe Data File](https://www.imls.gov/research-evaluation/data/museum-data-files).
  The 2018 file shows that a museum existed then, not that it exists now. This is a
  view of two lists, not a census: `r num(cp$source_rows)` records became
  `r num(cp$counted_institutions)` counted institutions.
- **Same name.** Names are compared after light tidying: case, punctuation, a leading
  "The", "&" for "and", and common abbreviations. Place names stay, because for a
  museum the place is usually the identity.
- **One museum, one vote.** Records for the same institution are merged first. A
  test of the merging step on 300 hand-labelled pairs is
  [written up here](https://github.com/jpbranson/duplicate-names/blob/main/data/validation/resolution_validation_2026-09-15.md);
  it measures those pairs, not the whole dataset.
- **Chains apart.** Museums with a known shared operator or brand
  (`r num(cp$chain_locations)` locations so far) are counted separately. For
  everything else, shared ownership is unknown rather than ruled out.
- **What I checked by hand.** Every record behind the names in this post: the
  `r word(top$museums)` Old Jail Museums and the `r word(nrow(records))` records in
  the `r word(nrow(groups))` collisions, between September 15 and October 5, 2026.
  The links in the tables and files are the evidence. Some of the
  `r word(top$museums)` carry a longer formal title as well: Tennessee's tourism
  office lists Winchester's as the Franklin County Old Jail Museum. Only
  `r cp$complete_factual_reviews` institutions in the whole dataset have had a full
  review, so every other number here is provisional, and I have not yet corrected
  the stray records above in the data.

The numbers behind this post:
[name counts](payload/name_ladder.csv),
[the `r word(top$museums)` Old Jail Museums](payload/headline_members.csv),
[the `r top$runner_up_names`-way tie](payload/runner_up_names.csv),
[every "old jail" name](payload/headline_related_names.csv),
[the `r word(nrow(groups))` collisions](payload/collision_groups.csv) with
[each record and its evidence](payload/collision_members.csv), and
[chains](payload/chains.csv). The pipeline is on
[GitHub](https://github.com/jpbranson/duplicate-names).
