---
title: "Who checks the museum names?"
subtitle: "A search for duplicate names becomes a test of what counts as a museum."
author: "J.P. Branson"
date: 2026-09-26
slug: duplicate-museum-names
categories: [data]
tags: [names, museums, maps]
draft: true
output:
  blogdown::html_page:
    toc: false
    fig_caption: true
---

```{r setup, include=FALSE}
dn_root <- "."
source("_setup.R", local = TRUE)
knitr::opts_chunk$set(fig.path = "figures/")
```

> **Working draft — unpublished.** The national ranking and collision reviews are
> unfinished. The numbers below describe documented changes to a provisional
> dataset; they do not identify the most duplicated museum name in America.

The International Cryptozoology Museum in Bangor, Maine, calls itself
[“The World's Only Cryptozoology Museum”](https://cryptozoologymuseum.com/).
That is a wonderfully testable-sounding claim. First, though, you need to decide
what would count as another museum.

Even this opening example leaves a trail. The museum's
[current visitor page](https://cryptozoologymuseum.com/plan-your-visit/) gives 490
Broadway in Bangor and says its former Portland and Hammond Street locations are
closed. Its home page dates the Bangor opening to June 1, 2026. A directory that
retains the old addresses can make one institution look like several.

Our source data contain four records for this institution. After resolution, one
current institution counts; the older records remain visible with exclusion reasons.
That finding checks the identity of these records. It does not prove a worldwide
uniqueness claim, especially in a project whose geographic scope is the United States.

## A name is not an institution

The starting ingredients are Overture Maps' August 19, 2026 Places release and
the IMLS 2018 museum archive. Their dates matter. The second source tells us about
an earlier museum universe; a current retrieval date cannot make it a current census.

The analysis preserves every original record and groups records into institutions.
A separate, explicit review layer can correct a mistaken group, reconcile a move,
or keep a contradictory record out of the count. The old record is not erased.
A history society's museum, office and mailing address can appear under several
names. Two buildings can house one institution. One society can also operate two
separately named museums.

At this checkpoint, the `r format(dn_payload("checkpoint.csv")$original_source_rows, big.mark=",")`
source rows yield `r format(dn_payload("checkpoint.csv")$provisional_institutions, big.mark=",")`
provisionally counted institutions, of which
`r format(dn_payload("checkpoint.csv")$eligible_l2, big.mark=",")` are eligible for the
name analysis. Only `r dn_payload("checkpoint.csv")$complete_factual_reviews`
institution descriptions have complete factual reviews, including sourced exclusions.
The remainder have pending or unstarted checks.

The first apparent leaders made that distinction unavoidable. Three county-society
names began this review batch at eleven institutions each. Supported corrections
reduced each to six. The remaining records still need review.

The next pass found the same problem in Wayne County: ten candidate institutions sharing the
society name became one unresolved bare-name candidate after duplicate institutions,
public museum names and shared operators were accounted for. The ten entries called
*Depot Museum* now have sourced, more specific public names. This does not settle
their entire factual review, and it does not establish a national winner.

```{r review-counts, fig.width=6, fig.asp=1.2, fig.alt="Before and after candidate counts: Franklin, Greene and Jackson County Historical Society each 11 to 6; Old Jail Museum 11 to 8; Wayne County Historical Society 10 to 1; Depot Museum 10 to 0.", fig.cap="Provisional L2 non-chain candidate counts before and after the September 26 review batches. Remaining factual reviews prevent a national ranking claim."}
dat <- dn_payload("review_count_changes.csv")
dat$group <- factor(dat$group, levels = rev(unique(dat$group)))
dat$checkpoint <- factor(dat$checkpoint, levels = c("Before review", "Current provisional"))
ggplot(dat, aes(count, group, fill = checkpoint)) +
  scale_y_discrete(labels = function(x) sub(" County Historical Society", paste(" County", "Historical Society", sep = intToUtf8(10)), x, fixed = TRUE)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65) +
  geom_text(aes(label = count), position = position_dodge(width = 0.75),
            hjust = -0.25, size = 4.1, colour = dn_ink$primary) +
  theme_dupnames(base_size = 14) + scale_fill_dn() + scale_x_continuous(limits = c(0, 12), breaks = c(0, 3, 6, 9, 12)) +
  labs(x = "Candidate institutions", y = NULL,
       title = "Review changes the apparent leaders",
       subtitle = "These are unresolved candidate groups, not certified winners.",
       caption = paste(dn_caption(c("overture", "imls")), "L2 names; affiliated museums separated.", sep = intToUtf8(10)))
```

Old Jail Museum illustrates several different corrections. Smethport's operator
uses [The County Museum in The Old Jail](https://www.mchsmuseum.org/) as its public
name. That changes its exact-name group without changing the source record.
Barnesville's [society](https://www.blchs.com/) distinguishes its jail museum from
its research archive. Allegan's [operator](https://alleganoldjailmuseum.com/) runs
both a jail museum and a historical village. These are three different reasons a
simple name count can mislead.

In Enterprise, Alabama, the [current operator](https://peariverhistorical.square.site/)
calls its attraction **Pea River Museum** and distinguishes it from its gift shop
and research library. In Corydon, Iowa, the
[Prairie Trails Museum contact page](https://prairietrailsmuseum.wordpress.com/contactus/)
connects a physical museum address with its post-office box. These corrections
preserve the source labels as aliases while counting the reviewed institution once.

The following groups began later review passes with nine candidates each. The
same combination of old addresses, broader society records, specific public names
and common operators changed these counts too. Every number remains provisional.

```{r recent-review-counts, results='asis'}
recent <- dn_payload("recent_review_counts.csv")
names(recent) <- c("Source name group", "Before review", "Current candidates")
knitr::kable(recent, row.names = FALSE, align = c("l", "r", "r"))
```

One particularly revealing case is Carroll County. The Missouri society's legal
record had a transposed street number; an exact EIN match in the IRS file connected
it to the museum's current address. In Georgia, one automatic cluster instead
contained two distinct organizations: a historical society and a genealogical
society. The latter's own [description](https://ccgsga.org/about/) and
[library resources](https://ccgsga.org/resources/) establish a research and
publishing role. The correction separates them before excluding that organization
from the museum count. The historical society's current museum review remains open.

Clinton County exposed a different source problem. Several archived records mix
the physical address of a museum in one state with the mailing address, website or
legal identifier of a society in another. For example, the Pennsylvania operator
places [Heisey House Museum at 362 East Water Street](https://www.clintonpahistory.org/),
while the Michigan operator places its
[museum at 106 Maple](https://sites.google.com/cchsmichigan.org/welcome).
Those street addresses also appear in contradictory IMLS records assigned to other
states. The review keeps each mixed record visible and uncounted, with its own
identifier, so it cannot contribute disputed aliases to an accepted museum.

## Repeated names, repeated operators

For the principal question, affiliated museums are reported separately from
independently named institutions. Unknown affiliation stays unknown.

The distinction includes local operators. An organization can have its own
independent nonprofit board and still operate several museums that share that
board. It also includes brands: two records labelled Madame Tussaud's Wax Museum
in Las Vegas turned out to describe one attraction at the
[operator's published address](https://lasvegas-support.madametussauds.com/hc/en-us/articles/115002250572-What-is-your-address).
Its current public name is Madame Tussauds Las Vegas.

The Smithsonian example contains a different trap. Its own
[visitor directory](https://www.si.edu/visit/museums) distinguishes its museums from
the separate Affiliates program. Some generic Smithsonian Institution records refer
to museums under that parent; two others match office and research locations that
the [Archives of American Art](https://www.aaa.si.edu/sites/default/files/Documents/2023-Annual-Report.pdf)
distinguishes from its exhibition gallery. Counting parent labels would confuse
organizational structure with independent naming.

## International what?

A search for *National*, *International*, *World* or *American* produces candidates
for a second question: do names suggesting broad scope collide?

A word search cannot answer that on its own. *African American* identifies a cultural
subject. *American Civil War* identifies a historical subject. *Old World* is an
idiom. A museum of international folk art can describe the range of its collection.
Even a genuinely institutional scope word does not automatically promise exclusivity.
The draft therefore makes no numerical claim about independent collisions among
singularity-claiming names. Those identities, affiliations and meanings still need
review together.

The first twelve leading scope-word groups illustrate the problem. The
[Seneca-Iroquois National Museum](https://senecamuseum.org/about/) is operated by
the Seneca Nation; interpreting its name as a claim to represent the entire United
States would misread it. A [county account of its 2018 move](https://historicpath.com/current-event/new-seneca-iroquois-national-museum-and-cultural-center-opens-salamanca-708)
also explains why old and current street addresses appeared as separate records.

Other apparent collisions are support facilities. A Smithsonian
[conservation contract](https://nmaahc.si.edu/sites/default/files/downloads/nmaahc_digitalarchivesspecialist_package.pdf)
distinguishes the African American history museum on Constitution Avenue from
its conservation labs on Maryland Avenue. The
[Old World Foundation](https://oldworldfoundation.org/about-owf.html) describes its
fundraising role for the Wisconsin museum. Neither a laboratory nor a fundraising
office supplies evidence of another public museum with the same name.

These corrections leave specific open questions. They do not turn a word search
into a validated count of institutions making competing scope claims.

For the generic tail, the analysis distinguishes exact names from subjects and
naming templates. Children's museums and county-history museums can resemble one
another linguistically without sharing an operator. A subject extracted from a name
is a clue, not a complete description of a collection.

## The method behind a future ranking

Museum headlines use **L2**: cleaned names with abbreviations expanded, while
retaining geography. **L3** removes geographic wording and answers a different
question. Each reviewed institution gets one canonical name and one vote. Original
names, aliases, source points, decisions and excluded records remain auditable.
A preferred-name correction changes the analysis name, not the original source text.

An independent human sample of 300 nearby candidate pairs supported keeping the
matching threshold at 0.85. Its unweighted pair-level precision was 99.17% and recall
93.70%. The sample was balanced across five similarity bands; these are not estimates
of dataset-wide accuracy and do not validate all final clusters or distant moves.
The [validation report](evidence/matching-validation.html) explains the design.

Visitor maps require another check. Publication coordinates are kept in separate
fields, with operator evidence and dated access wording. A corrected point does not
certify that a building is open. UC San Diego's
[Mandeville visit page](https://mandevilleartgallery.ucsd.edu/visit/index.html), for
example, still lists an indefinite closure when checked on September 26.

The remaining work is substantive: finish the leading-name and relevant multi-site
reviews, settle uncertain museum roles and affiliations, validate M2 meanings, and
then rerun the explicit publication gates. Until that happens, the interesting
result is how easily a directory can manufacture a collision.

The next post turns to churches: the repeated *First*, the spacing between them,
and how far the ordinal ladder goes. It will need its own source checks and
independent matching labels; the museum sample cannot validate it.

[Download the review-count figure data](payload/review_count_changes.csv).
[Read the current publication blockers](payload/publication_blockers.csv).
[Download the twelve-group scope-word review](payload/scope_word_review.csv).
The bundle is self-contained: knitting reads its CSV payloads and local helpers,
with no analysis checkout, database query or source download.





