---
title: "Post title"
subtitle: ""
author: "J.P. Branson"
date: 2026-01-01
slug: post-slug
categories:
  - data
tags:
  - names
  - maps
draft: true
output:
  blogdown::html_page:
    toc: false
    fig_caption: true
---

```{r setup, include = FALSE}
# Development template: dn_root points at the analysis repo. Before publishing,
# package _setup.R and its config/theme/embed dependencies with the bundle,
# adjust this source path, and verify a knit without the analysis checkout.
# Data chunks must read only the bundle's small payload/ files.
dn_root <- Sys.getenv("DUPNAMES_ROOT", unset = "../../..")
source(file.path(dn_root, "posts", "_setup.R"))
```

Opening paragraph.

## A chart

```{r a-chart}
# dat <- dn_payload("something.csv")
# ggplot(dat, aes(x, y)) +
#   geom_col(fill = dn_palette[["blue"]], width = 0.7) +
#   labs(title = "...", subtitle = "...", caption = dn_caption(c("overture", "imls")))
```

## A map

Built by the pipeline, not here — this chunk only emits the iframe.

```{r a-map, results = "asis"}
# cat(dn_iframe(
#   slug    = "territory-first-baptist",
#   caption = "Each point is a First Baptist Church.",
#   title   = "Map of First Baptist Church locations"
# ))
```

## Notes on method

Replace this guidance with a paragraph stating the source snapshots, counting unit,
normalization level, exclusions, affiliation rules, and manual verification behind
the headline. Museum name headlines use L2 (`name_expanded`); L3 comparisons strip
geography and answer a different question. Explain the comparison levels explicitly.

Export museum name-analysis payloads from `museum_analysis`, which uses the curated
`museum_records` layer; `entities` remains the automatic baseline. Explain any count
changes from source-specific identity corrections and link their dated evidence/audit.
Retain source aliases and distinguish former sites from current counting locations.
Keep contradictory `source_conflict` rows in the audit and exclude their disputed aliases
from accepted institutions. The review's `lon`, `lat` and `map_url` remain source points,
not certified visitor coordinates. Apply sourced map corrections in separate publication
fields with evidence, and verify access wording before presenting a visitor map.

Report resolution validation with its sampling limits. The September 15 museum
sample has 300 independently labelled candidate pairs within 150 m, equally sampled
across five similarity bands: precision 99.17%, recall 93.70%, and sample error 3.00%
at threshold 0.85. These unweighted results do not establish dataset-wide accuracy
or validate final clusters and multi-site merges. Link the archived validation report
when drafting the museum post; use the relevant evidence for later posts.

Before publishing the museum post, check category decisions affecting headlines, distinguish
chains from independent naming templates, and verify the top-20 collisions and
relevant multi-site cases. Show alternate names where entities have been merged.
An institution's `verified` status means factual identity/count review is complete;
assistant source research does not supply independent labels for measuring accuracy.
Keep unknown affiliation and unreviewed cases explicit.
Call `dn_assert_museum_publication_ready()` for the selected L2 headlines before exporting
payloads; the pipeline does not invoke this helper automatically. Its status checks do not
replace evidence review or the semantic check on M2 scope words. Use the current queue and
checkpoint links in `data/validation/README.md` when preparing the export, then bundle the
needed evidence references with the post.
Use CSV payloads for a blog build without `arrow`, and verify that the completed
bundle knits without the analysis checkout or a pipeline run.
