---
title: "Post title"
subtitle: ""
author: "J.P. Branson"
date: 2026-01-01
slug: post-slug
categories:
  - data
tags:
  - names
  - maps
draft: true
output:
  blogdown::html_page:
    toc: false
    fig_caption: true
---

```{r setup, include = FALSE}
# dn_root points at the analysis repo. When this bundle is copied into the
# blog, the payload/ directory travels with it and _setup.R comes along too,
# so nothing here reaches back into the pipeline.
dn_root <- Sys.getenv("DUPNAMES_ROOT", unset = "../../..")
source(file.path(dn_root, "posts", "_setup.R"))
```

Opening paragraph.

## A chart

```{r a-chart}
# dat <- dn_payload("something.csv")
# ggplot(dat, aes(x, y)) +
#   geom_col(fill = dn_palette[["blue"]], width = 0.7) +
#   labs(title = "...", subtitle = "...", caption = dn_caption(c("overture", "imls")))
```

## A map

Built by the pipeline, not here — this chunk only emits the iframe.

```{r a-map, results = "asis"}
# cat(dn_iframe(
#   slug    = "territory-first-baptist",
#   caption = "Each point is a First Baptist Church.",
#   title   = "Map of First Baptist Church locations"
# ))
```

## Notes on method

Every post carries a short, honest paragraph here: which normalization level the
counts use, what the sensitivity band looks like at the levels either side, and
what the entity-resolution error rate was. The numbers are artifacts of the
normalizer and the post should say so rather than implying the data speaks for
itself.
