---
title: "How many Firsts fit in one town?"
author: "J. P. Branson"
date: "2026-09-26"
draft: true
output: blogdown::html_page
---

```{r setup, include=FALSE}
dn_root <- "."
source("_setup.R", local = TRUE)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                     fig.width = dn_blog$fig_width, fig.asp = 0.8)
counts <- read.csv("payload/duplicate_counts.csv", stringsAsFactors = FALSE)
exclusivity <- read.csv("payload/census_place_exclusivity.csv", stringsAsFactors = FALSE)
territory <- read.csv("payload/territory_summary.csv", stringsAsFactors = FALSE)
```

<div class="draft-notice"><strong>Research draft — unpublished.</strong> These are
provisional estimates from automatically resolved Overture worship sites. Independent
matching and naming-style labels, leading-name identity reviews and final publication
checks are unfinished. No national ordinal record is certified.</div>

There is a [Twelfth Baptist Church in Boston](https://www.tbcboston.org/). The name
invites an apparently easy exercise: find the other eleven.

That exercise needs a time, a boundary and a definition of “other.” Congregations
move, merge and rename themselves. The current map cannot tell us which First a
Twelfth once followed. Nor does a numbered church necessarily count within a city.
[Sixteenth Tabernacle Beth El's own history](https://alexandria.cogasoc.org/history-of-the-tabernacle/)
identifies it as the sixteenth tabernacle chartered in Virginia. Looking for fifteen
predecessors inside Alexandria would begin with the wrong boundary. It also exposes
a category error: the [parent organization describes its faith as Judaism](https://www.cogasoc.org/).
The Alexandria and Miami tabernacles are therefore retained as sourced exclusions
from this draft's Christian-only statistics, despite the source's Christian category.

The museum question was whether a distinctive name actually singles out one
institution. Here the useful questions run the other way: how widely do churches
reuse names, and what geographic territory does a “First” occupy?

## A map of recorded names

This draft starts from Overture's August 19, 2026 release. Its worship categories
are more useful than a search for the word “religious”: that word's broad category
also contains organizations that may not be a worship location. The analysis keeps
those records as reviewable holds, along with reported permanent closures, missing
names and points outside the fifty states and DC.

Two other inputs answer different questions. GNIS supplies a frozen 2021 collection
of geographic church features, including explicitly historical names. The HIFLD
copy hosted by FEMA supplies legal names at geocoded tax addresses. Its geographic
extent does not cover Hawaii. Those records help investigate omissions and identity;
they do not independently certify present-day congregations. OSM samples are kept
as internal validation context and contribute no values to these published tables.

For name comparisons, the pipeline retains the original text, expands abbreviations,
and strips recognized place qualifiers at level L3. A blocklist protects ambiguous
words. “First Baptist Church of Peoria” and a comparable locative prefix can share a
core name; a shared core never establishes common ownership or one congregation.

```{r leading-names, fig.alt="Provisional L3 name counts. Removing geographic qualifiers makes Baptist Church the largest template; independent validation remains pending."}
leaders <- counts |> filter(level == "name_core") |> head(10)
ggplot(leaders, aes(x = n_entities, y = reorder(name_value, n_entities))) +
  geom_col(fill = dn_palette[["blue"]], width = .7) +
  labs(x = "Provisional worship-site candidates", y = NULL,
       title = "Names reused across the recorded sites",
       subtitle = "L3 comparison; identity and coverage validation pending") +
  theme_dupnames(base_size = 14)
```

Removing place qualifiers can turn many distinct public names into a broad template
like “Baptist Church.” The L3 chart counts those templates; the L2 payload preserves
the more literal comparison. Gazetteer ambiguity remains a validation concern.

The units here are provisional worship sites, with nearby automatic duplicates
consolidated. They are not a certified count of independent congregations. A church
with several campuses requires an institutional decision; no distant places are
merged merely because their names match. The payload also retains L2 and L4 results
so the effect of stripping place names or sorting tokens can be inspected.

## A First's nearest neighbor

A first description of territory measures the distance to the nearest **other** site
with the same L3 name. Distances follow the globe. A singleton has no observed neighbor;
it does not receive zero. Two different sites at the same point can receive zero,
which is a reason to review their identities rather than silently spread them apart.

<div class="dn-table-scroll" style="max-width:100%;overflow-x:auto;">

```{r distances}
shown <- territory |> filter(n_with_neighbor >= 30) |> arrange(desc(n_with_neighbor)) |> head(8)
shown$Family <- gsub("_", " ", shown$denom_norm)
knitr::kable(shown |> transmute(Family, `Sites with a neighbor` = n_with_neighbor,
  `Median km` = round(median_km, 1), `10th percentile km` = round(p10_km, 1),
  `90th percentile km` = round(p90_km, 1)), row.names = FALSE)
```

</div>

These distributions pool the observed First-name classes within broad denomination
families. Source omissions and false matches affect the distances. “Baptist” does
not identify a particular convention, and a name-derived family is a heuristic.
Conflicting category and name signals remain visible and are withheld from the
family-specific estimates.

<iframe data-external="1" title="Provisional First Baptist worship sites" src="embeds/first-baptist.html"
        width="100%" height="560" loading="lazy" style="border:0"></iframe>

[Open the interactive map](embeds/first-baptist.html) or view the [static fallback](first-baptist-fallback.png).

The map is a research view of source points. It is not a visitor guide and does not
certify current operation, institutional independence or location accuracy.

## One per place?

Spacing alone cannot tell us whether the convention is “one First per town.” The
second calculation asks: among **all Census places with any observed congregation
of a denomination**, how many contain exactly one observed First of that denomination?
Places containing a Second but no First remain in the denominator. Places with two
Firsts remain a separate outcome.

```{r place-pattern, fig.asp=0.85, fig.alt="Provisional shares of occupied Census places with zero, one, or multiple observed First churches, by broad denomination family."}
shown <- exclusivity |> arrange(desc(n_occupied_places)) |> head(8)
long <- tidyr::pivot_longer(shown, c(n_zero_first, n_exactly_one_first, n_multiple_first),
                           names_to = "pattern", values_to = "n")
long$pattern <- factor(long$pattern,
  levels = c("n_zero_first", "n_exactly_one_first", "n_multiple_first"),
  labels = c("No First", "One First", "Two or more"))
long$family <- gsub("_", " ", long$denom_norm)
ggplot(long, aes(x = n / n_occupied_places, y = reorder(family, n_occupied_places), fill = pattern)) +
  geom_col(width = .7) +
  scale_x_continuous(labels = function(x) paste0(round(100*x), "%")) +
  scale_fill_dn() +
  labs(x = "Share of occupied Census places", y = NULL, fill = NULL,
       title = "The denominator includes places without a First",
       subtitle = "Provisional observations, not a test of historical intent") +
  theme_dupnames(base_size = 14) + theme(legend.position = "bottom")
```

Census places include incorporated places and census-designated places. They do not
cover every settlement or all rural territory, and a CDP is not a municipality. The
bundle includes a separate incorporated-place calculation and the coverage audit.
No point outside a place is assigned to the nearest city for convenience.

Even a high observed rate would not establish a universal naming rule. The relevant
historical territory might be a state, a neighborhood or a denomination's own district.
The Sixteenth Tabernacle example makes that qualification concrete.

## Which numbers count?

The largest apparent ordinal is especially vulnerable to a small error. In the
initial extract, “107th Psalm,” “21st Century,” numbered intersections and administrative
districts all appeared among high-number candidates. Their numbers are not counts of
congregations. Two source records called “Thirty Eighth Baptist Church” linked to
[38th Avenue Baptist Church](https://38thavenuebaptist.org/contact-us): the missing
word **Avenue** was doing almost all the analytical damage.

The parser now protects those patterns and retains sourced public-name corrections
separately from raw records. It also holds out day expressions: [Seventh Day Baptists](https://seventhdaybaptist.org/wp-content/uploads/2016/03/Statement-of-Belief-QA-Tract.pdf)
refer to the Sabbath, while [Eighth Day Church explains its name in terms of new birth](https://www.eighthdaychurch.com/about-edc).
An outer number such as **First** in “First Seventh Day Baptist” still counts as a
candidate congregation ordinal.

Some histories do explain a sequence. [Tenth Presbyterian's history](https://www.tenth.org/about/history/)
identifies it as Philadelphia's tenth Presbyterian church. By contrast,
[Tenth Memorial Baptist](https://tenthmemorial.org/history/) commemorates the members
of an earlier Tenth Baptist congregation. These histories make the interpretation
more precise; they do not certify a present-day ladder of ten active congregations.
The remaining high ordinals still need identity and meaning checks. This draft
therefore names no national maximum.

Within a place and denomination, a provisional ladder records the distinct observed
rungs up to the largest parsed number. An absent Third could reflect missing data,
a name change, a move, a different historical boundary or a congregation that never
existed. The name list does not choose among those stories.

## Counting, saints, virtues and brands

The final comparison classifies names as ordinal, saint, virtue, toponym, modern brand,
ethnolinguistic, descriptive or other. This is currently a lexicon-based heuristic.
An independent sample of 500 names for estimating its errors is archived separately from
its predictions; it has not been graded by this analysis. Until that review is complete,
the proportions are research outputs, not a supported claim about denominational culture.

The pipeline also emits places with multiple Firsts as a distinct research file.
Explaining those cases requires historical sources. A later post will examine that
history; another will ask how naming changed over time, using dated evidence that
these snapshots do not provide. Neither story can be recovered merely by sorting names.

The [museum draft](../duplicate-museum-names/) begins this series. Both drafts retain
explicit publication gates and build from their own small payloads without
the analysis checkout. The independent-review protocol accompanies
the research bundle; full acquisition provenance remains in the repository manifest.







